Land-use legacies and climate change as a double challenge to oak forest resilience: mismatches of geographical and ecological rear edges

### 2020-03-10 man/ecosystems/ms_ecosystems.Rmd
aspect / exposure

By using a combined approach of remote-sensing information and dendrochronology, we have quantified the drought impact on the *Quercus pyrenaica* forests of Sierra Nevada and their resilience to several severe drought events in the recent decades. Our results indicated that these relict oak populations driven by historical land use are resilient to climate change at their present rear edge. However, resistance, resilience, and forest recovery to extreme drought events are strongly influenced by mountain aspect, local environmental conditions, and management legacies. This means that the geographical and the ecological rear edges do not necessarily match and, at a small spatial scale, tree performance varies markedly along the rear edge under climate change.

>>>>

By using a combined approach of remote-sensing information and dendrochronology, we have quantified the drought impact on the *Quercus pyrenaica* forests of Sierra Nevada and their resilience to several severe drought events in the recent decades. Our results indicated that these relict oak populations driven by historical land use are resilient to climate change at their present rear edge. However, resistance, resilience, and forest recovery to extreme drought events are strongly influenced by mountain exposure, local environmental conditions, and management legacies. This means that the geographical and the ecological rear edges do not necessarily match and, at a small spatial scale, tree performance varies markedly along the rear edge under climate change.
